Automated Topic Naming

### 2012-01-19 ese/ese.tex
minor edit
latest greatest edits?
minor fixes and typos

%Researchers have employed a variety of techniques to extract underlying
%topics that relate to software development artifacts.  Typically, these
%techniques use semi-unsupervised machine-learning algorithms to suggest
%candidate word-lists, topics.  
Software repositories provide a deluge of software artifacts to
analyze. Researchers have attempted to summarize, categorize, and relate these
artifacts by using semi-unsupervised machine-learning
algorithms, such as Latent Dirichlet Allocation (LDA), used for concept and topic analysis to suggest candidate
word-lists or topics that describe and relate software artifacts.
However, these word-lists and topics are
difficult to 
interpret in the absence of meaningful summary labels.  Current topic modeling 
techniques assume manual labelling and do not use domain-specific knowledge
to improve, contextualize, or describe results for the developers.  
We propose a solution: \emph{automated labelled topic extraction}.
  Topics are extracted 
using  LDA from commit-log comments recovered
from source control systems. % such as CVS and BitKeeper.  
These topics are given labels from a generalizable
cross-project taxonomy, consisting of non-functional
requirements. Our approach was evaluated with experiments and case studies
on three large-scale Relational Database Management System (RDBMS) projects: MySQL, PostgreSQL and MaxDB.  
The case studies
show that labelled topic extraction can produce appropriate,
context-sensitive labels that are relevant to these projects, and provide fresh
insight into their evolving software development activities.
%160 words
\keywords{ Software maintenance \and Repository mining \and Latent Dirichlet allocation \and Topic models}

@@

Topic modeling (such as Latent Dirichlet Allocation ~\cite{Blei2003}) is a machine learning technique that creates
multinomial distributions of words extracted from a text corpus. 
This technique infers the hidden structure of a corpus using posterior
inference: the probability of the hidden structure given the data. 
Topic models are useful in software maintenance because they summarize
the key concepts in a corpus -- such as source code, commit comments, or
mailing-list messages -- by identifying words that commonly occur together. 
Among other uses, topic modelling can quickly give developers an overview of where significant
activity has occurred, and provide managers or maintainers an enhanced understanding of the project's 
history.

@@

Figure~\ref{fig:pgcomp} shows the ROC results for the PostgreSQL
product. In this figure, we show the relative differences when we use
different training data sets from different annotators. In \ref{fig:pgcomp-a} we see that ROC
results are very similar. In the other figure, however, author \#1 has
dramatically better performance. We speculate that this is due to the
particular annotation decisions made by author \#1; in some sense he
performed better. 
The difference in Macro-ROC was not significant, but the difference in
Micro-ROC was, as the p-value of Student's T-test was $< 0.001$. 
%\DONE{check this writeup of comparison. T-test useful}

@@

\subsection{RQ1: Do NFR Frequencies Change over Time?}
In both MaxDB and MySQL the frequencies generally decreased with age. 
However, there are variations within our NFR labels. In MySQL, \emph{usability} and \emph{efficiency} do not appear very often in topics. 
A proportionately smaller number of commits addressed these NFRs.
%As discussed in the previous section, 
Certain peaks in topic numbers coincide with a particular emphasis from the development team on issues such as new releases or bug fixes.
This suggests that maintenance activity is not necessarily strictly decreasing with time, but rather episodic and responsive to outside stimuli. 
In MaxDB, we can observe that \emph{Maintainability} topics became more prevalent as MaxDB matures. 
This is likely due to our analysis time-frame for MaxDB being shorter than the time-frame for the MySQL product.

@@

Based on Figure \label{fig:authorsim} we wanted to see which clusters
\texttt{momjian}, \texttt{tgl}, and \texttt{petere} would be in if we
evaluated the clusters proportionately, that is if we use distance
measures that normalize the data and ignore magnitude (e.g. turning an
author NFR contributions into a histogram or unit-vector) such as
cosine distance, euclidean distance of unit vectors, or Pearson
correlation. Euclidean distance uses both angle and magnitude of a
vector. For all 3 non-magnitude distances measures we tend to see that
the 3 largest contributors, \texttt{momjian}, \texttt{tgl}, and
\texttt{petere} inhabit the same clusters when we use 2 clusters or 6
clusters. This confirms the observations in
Figure \label{fig:authorsim}, which uses Pearson correlation, that the larger contributors tend to be
proportionally similar.

@@

\subsection{Annotation Observations}
We found many topics that were not non-functional requirements (NFRs) but were often related to them. 
For instance, concurrency was mentioned often in the commit logs and
was related to \emph{correctness} and \emph{reliability}, possibly because concurrent code is prone to bugs such as race conditions. %are 
% often  was troublesome. 
Topics related to \emph{Configuration management} and \emph{source control} appeared often; % and sometimes there were topics dedicated to configuration
% management. 
these kinds of changes are slightly related to \emph{maintainability}. 
A non-functional change that was not quality-related was licensing and
copyright; many changes concerned updating copyrights or
ensuring copyright or license headers were applied to files. In these
cases we assigned the \emph{None} label to the topic.

@@

We feel these results are poor. The aggregate view of a Kappa of
$0.1$ indicates there is some weak agreement. We found that
there was a lot of agreement in terms of lack of a tag, but often more
disagreement when a tag should be applied. After some discussion we concluded that 
usability was a primary source of
disagreement. Do user manuals improve usability? Is a command-line
option a usability issue? These kinds of questions illustrate some of
the agreement, disagreement, and ambiguity about these labels. 
We therefore recommend that future annotators train and discuss how and when an annotation is appropriate.

To evaluate our results empirically, we compared them the annotations that would result if the seven labels
 were applied randomly. In the first simulation
we sampled with replacement from our own NFR label distributions to
produce random ratings that looked like our own. We then applied the
Kappa-statistic on this sampling, compared it with our labelling, and repeated
this 100,000 times. We found that
for 4 out of 7 NFR labels, the IRR values for our labels were greater than the random
IRR measures 96\% of the time. For \emph{portability} and \emph{efficiency}, our IRR
was greater 100\% of the time both for Neil's ratings, Abram's ratings
and the union of both Neil and Abram. Figure \ref{fig:irr} depicts the IRR of
these sampled ratings versus our IRR using the Kappa statistic. We also achieved similar results if the ratings were pulled from a
uniform distributions. These experiments indicate that our labels and
ratings were an improvement on random annotations in all cases, with the exceptions of \emph{functionality},
\emph{reliability} and \emph{usability}.

\begin{figure}
  \centering
  \includegraphics[width=0.9\textwidth]{figures/self-sample-Both}
  \caption{Measured IRR versus IRR of labeling random simulations.}
  \label{fig:irr}
\end{figure}

Were our results harmed by low IRR scores? For MaxDB and MySQL since
each annotator (Neil or Abram) acted as the sole Oracle, the results
for either system still stand, but construct validity is harmed in
terms of did if an NFR actually meant the NFR that is being
evaluated. The measured empirical performance of each system still
stands, but the comparison between systems in terms of NFRs might be
harmed by low IRR scores.

We feel that this resulting IRR score indicates
the underlying difficulty of annotating such a data-set. To improve IRR,
consistent training of the annotators or the use of more annotators might help.
Other improvements could be gained by approaching some of the original developers
and using their expertise to annotate the data according to their original, recollected intent.
Improvements
could be therefore be gained by producing more robust training samples that use
multiple raters, or experts, who are similarly trained and willing to discuss and
negotiate disagreements.

>>>>

%Researchers have employed a variety of techniques to extract underlying
%topics that relate to software development artifacts.  Typically, these
%techniques use semi-unsupervised machine-learning algorithms to suggest
%candidate word-lists, topics.  
Software repositories provide a deluge of software artifacts to
analyze. Researchers have attempted to summarize, categorize, and relate these
artifacts by using semi-unsupervised machine-learning
algorithms, such as Latent Dirichlet Allocation (LDA), used for concept and topic analysis to suggest candidate
word-lists or topics that describe and relate software artifacts.
However, these word-lists and topics are
difficult to 
interpret in the absence of meaningful summary labels.  Current topic modeling 
techniques assume manual labelling and do not use domain-specific knowledge
to improve, contextualize, or describe results for the developers.  
We propose a solution: \emph{automated labelled topic extraction}.
  Topics are extracted 
using  LDA from commit-log comments recovered
from source control systems. % such as CVS and BitKeeper.  
These topics are given labels from a generalizable
cross-project taxonomy, consisting of non-functional
requirements. Our approach was evaluated with experiments and case studies
on three large-scale Relational Database Management System (RDBMS) projects: MySQL, PostgreSQL and MaxDB.  
The case studies
show that labelled topic extraction can produce appropriate,
context-sensitive labels that are relevant to these projects, and provide fresh
insight into their evolving software development activities.
%173 words
\keywords{ Software maintenance \and Repository mining \and Latent Dirichlet allocation \and Topic models}

@@

Topic modeling (such as Latent Dirichlet Allocation~\cite{Blei2003}) is a machine learning technique that creates
multinomial distributions of words extracted from a text corpus. 
This technique infers the hidden structure of a corpus using posterior
inference: the probability of the hidden structure given the data. 
Topic models are useful in software maintenance because they summarize
the key concepts in a corpus -- such as source code, commit comments, or
mailing-list messages -- by identifying words that commonly occur together. 
Among other uses, topic modelling can quickly give developers an overview of where significant
activity has occurred, and provide managers or maintainers an enhanced understanding of the project's 
history.

@@

Figure~\ref{fig:pgcomp} shows the ROC results for the PostgreSQL
product. In this figure, we show the relative differences when we use
different training data-sets from different annotators. In \ref{fig:pgcomp-a} we see that ROC
results are very similar. In the other figure, however, author \#1 has
dramatically better performance. We speculate that this is due to the
particular annotation decisions made by author \#1; in some sense he
performed better. 
The difference in Macro-ROC was not significant, but the difference in
Micro-ROC was, as the p-value of Student's T-test was $< 0.001$. 
%\DONE{check this writeup of comparison. T-test useful}

@@

\subsection{RQ1: Do NFR Frequencies Change Over Time?}
In both MaxDB and MySQL the frequencies generally decreased with age. 
However, there are variations within our NFR labels. In MySQL, \emph{usability} and \emph{efficiency} do not appear very often in topics. 
A proportionately smaller number of commits addressed these NFRs.
%As discussed in the previous section, 
Certain peaks in topic numbers coincide with a particular emphasis from the development team on issues such as new releases or bug fixes.
This suggests that maintenance activity is not necessarily strictly decreasing with time, but rather episodic and responsive to outside stimuli. 
In MaxDB, we can observe that \emph{Maintainability} topics became more prevalent as MaxDB matures. 
This is likely due to our analysis time-frame for MaxDB being shorter than the time-frame for the MySQL product.

@@

Based on Figure \ref{fig:authorsim} we wanted to see which clusters
\texttt{momjian}, \texttt{tgl}, and \texttt{petere} would be in if we
evaluated the clusters proportionately, that is if we use distance
measures that normalize the data and ignore magnitude (e.g. turning an
author's NFR contributions into a histogram or unit-vector) such as
cosine distance, Euclidean distance of unit vectors, or Pearson
correlation. Euclidean distance uses both angle and magnitude of a
vector. For all three non-magnitude distances measures we tend to see that
the three largest contributors, \texttt{momjian}, \texttt{tgl}, and
\texttt{petere} inhabit the same clusters when we use 2-clusters or 6-clusters. This confirms the observations in
Figure \ref{fig:authorsim}, which uses Pearson correlation, that the larger contributors tend to be
proportionally similar.

@@

\subsection{Annotation Observations}
We found many topics that were not non-functional requirements (NFRs) but were often related to them. 
For instance, concurrency was mentioned often in the commit logs and
was related to \emph{correctness} and \emph{reliability}, possibly because concurrent code is prone to bugs such as race conditions. %are 
% often  was troublesome. 
Topics related to \emph{configuration management} and \emph{source control} appeared often; % and sometimes there were topics dedicated to configuration
% management. 
these kinds of changes are slightly related to \emph{maintainability}. 
A non-functional change that was not quality-related was licensing and
copyright; many changes concerned updating copyrights or
ensuring copyright or license headers were applied to files. In these
cases we assigned the \emph{None} label to the topic.

@@

These results are fairly poor. The aggregate view of a Kappa of
$0.1$ indicates there is some weak agreement. We found that
there was good agreement in terms of lack of an annotation, but
disagreement regarding which annotation to apply. After some discussion we concluded that 
\emph{usability} was a primary source of
disagreement. For instance, should we annotate a commit which updates the user manual as a usability-related change? Is adding a command-line
option a usability issue? These kinds of questions illustrate some of
the agreement, disagreement, and ambiguity about these labels. 
We therefore recommend that future annotators train and discuss how and when an annotation is appropriate. This ought to be
easier to do if an internal team is using our tool, since (presumably) there is a greater shared understanding of project activities, and therefore,
potentially greater rater agreement.

To evaluate our results empirically, we compared them to the annotations that would result if the seven labels
 were applied randomly. In the first simulation
we sampled with replacement from our own NFR label distributions to
produce random ratings that looked like our own. We then applied the
Kappa-statistic on this sampling, compared it with our labelling, and repeated
this 100,000 times. We found that
for 4 out of 7 NFR labels, the IRR values for our labels were greater than the random
IRR measures 96\% of the time. For \emph{portability} and \emph{efficiency}, our IRR
was greater 100\% of the time both for Neil's ratings, Abram's ratings
and the union of both Neil and Abram. One explanation of the positive result for \emph{portability} is that such changes are often 
accompanied by clear indicator terms such as \emph{Windows, Linux, SunOS}, etc.

Figure \ref{fig:irr} depicts the IRR of
these sampled ratings versus our IRR using the Kappa statistic. We also achieved similar results if the ratings were pulled from
uniform distributions. These experiments indicate that our labels and
ratings were an improvement on random annotations in all cases, with the exceptions of \emph{functionality},
\emph{reliability} and \emph{usability}.

\begin{figure}
  \centering
  \includegraphics[width=0.9\textwidth]{figures/self-sample-Both}
  \caption{Measured IRR versus IRR of labeling random simulations. Perfect IRR is 1.0. }
  \label{fig:irr}
\end{figure}

Were our results harmed by low IRR scores? For MaxDB and MySQL, since
each annotator (Neil or Abram) acted as the sole oracle, the results
for either system still stand, but construct validity is harmed since 
it is not clear if the annotated NFR actually represented the NFR relevant to that commit.
 The measured empirical performance of each system still
stands, but the comparison between systems in terms of NFRs might be
harmed by low IRR scores.

The low IRR scores indicate
the underlying difficulty of annotating such a data-set. To improve IRR,
consistent training of the annotators or the use of more annotators might help.
Other improvements could be gained by approaching some of the original developers
and using their expertise to annotate the data according to their original, recollected intent.
Improvements
could be therefore be gained by producing more robust training samples that use
multiple raters, or experts, who are similarly trained and willing to discuss and
negotiate disagreements.
